---
title: "Project1"
author: "arnela82"
date: "4/29/2020"
output: html_document
---
##This is my submission for Reproducible Research Course Project 1

###Loading and preprocessing the data
```{r, echo=TRUE}
activity<-read.csv("C:/Users/Arnela/Desktop/coursera/activity.csv")
summary(activity)

```

```{r,echo=TRUE}
#Some information about the variables
str(activity)
```

```{r,echo=TRUE}
#Some information about the variables
names(activity)
```

```{r,echo=TRUE}
#Some information about the variables
head(activity)
```

###Total number of steps per day
*1.We calculate the total number of steps taken per day*
```{r echo=TRUE}
StepsPerDay<-aggregate(steps~date,data = activity,sum, na.rm=TRUE)
colnames(StepsPerDay)<-c("Date","Steps")
StepsPerDay
```

*2.Histogram of the total number od steps taken each day*
```{r histogram1, echo=TRUE, fig.width=10, warning=FALSE}
hist(StepsPerDay$Steps)
```

*3.Mean and median of total number of steps taken per day*
```{r echo=TRUE}
#mean
mean(StepsPerDay$Steps,na.rm = TRUE)
#median
median(StepsPerDay$Steps,na.rm = TRUE)
```

<br>
###Averagre daily activity pattern

*1.Time series plot of the 5 minute interval and averaged number of steps taken averaged across all days*
```{r timeplot1,echo=TRUE, fig.width=10}
#create table with steps per time
StepsPerInterval<-aggregate(steps~interval,data = activity,mean,na.rm=TRUE)
#variable time
StepsPerInterval$time<-StepsPerInterval$interval/100
plot(steps~interval,data = StepsPerInterval)

```

*2.5-minute interval with the max number of steps*
```{r echo=TRUE}
#The 5-minute interval accross all day containig the maximum nuber od steps is stored in variable MaxNbSteps
MaxNbSteps<-StepsPerInterval[which.max(StepsPerInterval$steps),]$interval
MaxNbSteps

```
<br>

###Imputing missing values
*1.Total number od missig values ( totalMV )*
```{r echo=TRUE}
totalVM<-sum(is.na(activity$steps))
totalVM
```

*2.Replacing missing values*
```{r echo=TRUE}
getMeanStepsPerInterval<-function(interval){
  StepsPerInterval[StepsPerInterval$interval==interval,]$steps
}
```

*3.Create a new dataset that is equal to the original dataset but with the missing data filled in*
```{r echo=TRUE}
#new dataset is activityFull
activityFull<-activity
for(i in 1:nrow(activityFull)){
  if(is.na(activityFull[i,]$steps)){
    activityFull[i,]$steps<-getMeanStepsPerInterval(activityFull[i,]$interval)}
}
```

*4.Histogram of the total number of steps taken each day with missing data filled in*
```{r histogram2,echo=TRUE, fig.width=10}
totalStepsPD<-aggregate(steps~date,data=activityFull,sum)
hist(totalStepsPD$steps)

```
*4.a Caltulate and report the mean and median total number of steps taken per day*
```{r echo=TRUE}
#mean
mean(totalStepsPD$steps)
#median
median(totalStepsPD$steps)
```

<br>

###Are there differences in activity patterns between wkkdays and weekends?
*1.Create a new factor variable in the dataset with two levels - weekday and weekend indicating whethever a given date is a weekday or weekend day.*
```{r echo=TRUE}
activityFull$date<-as.Date(activityFull$date, format = "%Y-%m-%d")
activityFull$day<-weekdays(activityFull$date)
for (i in 1:nrow(activityFull)) {
  if (activityFull[i,]$day %in% c("Saturday","Sunday")){
    activityFull[i,]$day<-"weekday"
  }
}
```

*2.Two time series plot of the 5-minute interval and the average number of steps accross weekday days and weekend days*
```{r timeplot2,echo=TRUE, fig.width=10}
stepsByday<-aggregate(activityFull$steps~activityFull$interval+activityFull$day, activityFull,mean)
names(stepsByday)<-c("interval","day","steps")
library(lattice)
xyplot(steps~interval|day, stepsByday, type="1",layout=c(1,2),xlab="Interval",ylab = "Number of steps")
```




